Recent Advancements in Neural Computation: \\ \large Differentiable Neural Computers and End-To-End Memory Networks

### 2018-06-23 Paper/Paper.tex
final

\documentclass[]{article}
\usepackage[a4paper]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathabx}
\usepackage{graphics}
%\usepackage{mathtools}
\usepackage{tikz}
\usepackage{array}
\usepackage{proof}
\usepackage{hyperref}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}
\usepackage{subcaption}
\usetikzlibrary{calc, trees, fit, positioning}

@@

A particularly interesting class of neural networks are \textit{Recurrent Neural Networks} (RNNs). Unlike fully-connected networks, which accept a fixed-size input vector and produce a fixed-size output vector over a finite number of computational steps, RNNs instead operate on sequences of vectors. This additional axis of dependencies is captured by a recurrency relation interlinking the network's hidden states through time. More concretely, the hidden representation is no longer conditioned on just the current input, but also on the immediately previous version of itself (Fig.\ref{fig:1}). This elegant addition accounts for a drastic improvement in the representational capacity of the network. Whereas a feedforward net acts as a trainable (but parametrically fixed) function, a RNN acts as a continuous program operating on sequences of inputs, and is in fact Turing complete\cite{Siegelmann545}. Under this viewpoint, the RNN's hidden representation can be perceived as the program's internal variables, updating as new inputs are received and affecting the computation as it progresses.

\begin{figure}
\begin{minipage}{0.5\textwidth}
	\centering
	\begin{tikzpicture}[scale=0.5]
		\node (X) at (-1, -2) {$\pmb{x}$};
		\foreach \y /\alph/\name in {
			0/x1/{},
		  	-1/x2/{},
		  	-2/x3/{},
		  	-4/xM/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=4mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (H) at (4, 2.5) {$\pmb{h}$};
		 \node (x..) at (0,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			1.5/h1/{},
		  	0/h2/{},
		  	-1.5/h3/{},
		  	-5.5/hN/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (4,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \node (Y) at (9, -2) {$\pmb{y}$};
		 \foreach \y /\alph/\name in {
			0/y1/{},
		  	-1/y2/{},
		  	-2/y3/{},
		  	-4/yP/{}
		 }{
		 	\node[circle, draw=gray!90,inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
	\end{tikzpicture}
	\subcaption{Simple Neural Net}
	\end{minipage}
	\begin{minipage}{0.5\textwidth}
	\centering
	\begin{tikzpicture}[scale=.5]
		\node (X) at (-1, -2) {$\pmb{x}$};
		\foreach \y /\alph/\name in {
			0/x1/{},
		  	-1/x2/{},
		  	-2/x3/{},
		  	-4/xM/{}
		 }{
		% 	\node[] (\alph) at (0,\y) {\name};
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (x..) at (0,-3) {\tiny {\vdots}};
	  \node (H) at (4, 2.5) {$\pmb{h}$};
		 \foreach \y /\alph/\name in {
			1.5/h1/{},
		  	0/h2/{},
		  	-1.5/h3/{},
		  	-6.5/hN/{}
		 }{
		 	\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=gray!20] (\alph) at (4,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \node (Y) at (9, -2) {$\pmb{y}$};
		 \foreach \y /\alph/\name in {
			0/y1/{},
		  	-1/y2/{},
		  	-2/y3/{},
		  	-4/yP/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(0.8,0.8) and +(-0.8,0.8) .. (h1.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(0.4,-0.8) and +(-0.4,0.8) .. (h2.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(1,-1.6) and +(-0.6,0.4) .. (h3.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(2,-2) and +(-2,2) .. (hN.west);
\end{tikzpicture}
	\subcaption{Recurrent Neural Network}
	\end{minipage}
	\caption[]
	{\small \textit{Comparison between a feedforward neural network and a RNN. (a) The simple network applies two transformations $\pmb{x} \mapsto \pmb{h}$ and $\pmb{h} \mapsto \pmb{y}$, the composition of which can be seen as a function $\pmb{x} \mapsto \pmb{y}$. (b) The recurrent network allows through-time connections, depicted with the dotted lines, such that $\pmb{x}_t \times \pmb{h}_{t-1} \mapsto \pmb{h}_t$. The output then becomes a function of the full sequence of inputs received: $\underbrace{\pmb{x}_0 \times \pmb{x}_1 \times \dots \times \pmb{x}_t}_{\text{Sequence Length}} \mapsto \pmb{y}_t$.}\label{fig:1}}
\end{figure}

@@

A series of advancements has seeked to enhance RNNs with stronger memory management. Starting with the introduction of the \textit{Long Short-Term Memory}\cite{Hochreiter:1997:LSM:1246443.1246450,Gers99learningto}, which permits finite storage over very long computations, research in RNNs has bloomed towards many different directions. One particularly exciting recent advancement is the so-called \textit{attentional interface}\cite{NIPS2014_5542,Bahdanau} (Fig.\ref{fig:2}), which allows a network to selectively shift its focus on particular indices over a set of dimensions of an arbitrary tensor, thus enabling content-based adressing. The implications of this seemingly simple trick are tremendous; when employing fully differentiable attention (soft-attention), a network can learn how to pick out and combine relevant elements out of a (potentially very large) tensor, instead of lossily encoding it. This translates to arbitrary information retrieval; in other words, the same representations can now have different effects on the computation depending on the present context, similar to how a computer program would refer only to the variables dictated by its currently run instruction\cite{NIPS2014_5346}. Since its inception, attention has been a key-concept in many influential publications; the creative liberty it is offering, in terms of new potential operations, has also paved the way for architectures which decouple storage from computation. Such architectures, including (but not limited to) the two presented in this paper, utilize large external matrices and employ attention to perform read and write operations on them.

@@

		\draw (Ny1)  [dashed] .. controls +(0.8,0.8) and +(-0.8,0.8) ..  (x1);
		\draw (Ny2)  [dashed] .. controls +(3,3) and +(-0.6,0.6) ..  (x2);
		\draw (Ny3)  [dashed] .. controls +(2,-3) and +(0.6,0.4) ..  (xT.west);
	\end{tikzpicture}
	\caption[]
	{\small \textit{\label{fig:2} General case of an attentional layer. Given some sequence $[\pmb{h}_1,\dots,\pmb{h}_T]$ (e.g. a full sequence of RNN hidden states), we may use an neural function $a$ to obtain a weighting vector $\pmb{a} \in \mathbb{R}^{T}$. Multiplying each sequence element by its corresponding weight and summing up the results, we end up with a context vector $\pmb{c}$. Note that when $a$ accepts an external vector (or vector sequence) as its argument, the result is a context-dynamic selection and linear combination of (parts of) the $[\pmb{h}_1,\dots]$ sequence.}}
\end{figure}

@@

\section{Differentiable Neural Computers}
\label{DNC}
\subsection{Description}
Inspired by the von Neumann architecture, Differentiable Neural Computers (or DNCs in short) model computation as the interfacing between two units; a RNN controller and an external memory matrix, functioning analogously to the CPU and the RAM, respectively. The memory matrix can be adressed using soft attention mechanisms, allowing selective reading and writing. As both the controller and the interfacing are differentiable, the full architecture is end-to-end differentiable, essentially allowing it to be trained purely with data and minimal human supervision. As the system is already equiped with the necessary tools to simulate traditional computation, i.e. adressing mechanisms, information storage and retrieval, logical flows and arithmetic processing, a DNC in training is learning the instructions of a program rather than the parameters of a function. A trained DNC should then be able to mimic the behavior of a traditionally hand-written program.

@@

The retention vector is a measure of how much each memory address needs to be retained. Its functional complement, the \textit{usage-tracking vector} $\pmb{u}_t$, annotates each memory address with a scalar, responsible for signaling how strongly it is being used. It is recurrently defined as follows:
\[
\tag{Usage-tracking Vector}
\pmb{u}_t = \underbrace{(\pmb{u}_{t-1} + \pmb{w}_{t}^{w} - \pmb{u}_{t-1} \circ \pmb{w}_t^{w})}_{\text{ \scriptsize Previous usage OR New write}} \circ \ \pmb{\psi}_t
\]
Contrary to first impressions, the above equation is quite intuitive. The parenthesized term sums the previous usage tracking vector with the current write weighting minus their dot product, acting as a real-valued variation of the logical OR function. A single element of this term, $\pmb{u}_t[i]$ would denote the pre-retention usage tracking of memory address $i$; it is close to $1$ if it was either already in use, or just written to. Multiplied by its corresponding element of the retention vector, it can then be (partially) deallocated.

@@

\paragraph{Temporal Linking} One last mechanism of memory addressing employed by DNCs is temporal linking. Temporal linking is necessary in order to perform sequential retrieval, when sequences are stored in disparate, non-contiguous blocks. The key-element of the process is a temporal transition matrix $\pmb{L} \in [0,1]^{N\times N}$, which keeps track of the temporal order by which memory addresses have been written to. More conretely, $\pmb{L}[i,j]$ is close to $1$ if $j$ was strongly written to after $i$, or close to $0$ otherwise. $\pmb{L}$ is again defined via a recurrency relation; a slightly simplified version of the one in the original paper is presented below, for the sake of brevity and clarity.
\[
\tag{Temporal Update}
\pmb{L}[i,j] = \underbrace{(1 - \pmb{w}_t^w[i] - \pmb{w}_t^w[j]) \pmb{L}_{t-1}[i,j]}_{\text{\scriptsize Part of last transition}} + \underbrace{\pmb{w}_{t-1}^w[i] \pmb{w}_{t}^w[j]}_{\text{\scriptsize New transition}}
\]
The above equation is also quite intuitive. The first term of the addition allows the transition value between $i$ and $j$ to propagate through time. The extend of propagation of the old transition, $\pmb{L}_{t-1}[i,j]$, is complentary to the amount of writing performed in these addresses; even if we had a transition in the past, it no longer holds if either $i$ or $j$ were just written to. On the other hand, if $i$ was written to in the previous timestep and $j$ was written to in the current timestep, then a new transition is instantiated. This is captured by the second term of the addition.

@@

\section{End-to-End Memory Networks}
\label{E2E}
\subsection{Description}
An end-to-end memory network is a neural architecture that also connects a RNN with a large memory array. The external nature of this array is disanalogous to traditional architectures, such as the LSTM, where the memory is encapsulated within the neural layers. Considering this difference, it can be perceived as true long-term memory. The memory can be read from (potentially multiple times) during computation, using a straightforward attentional layer; the architecture therefore benefits from context-sensitive addressing capabilities. The overall model is simple, yet efficient, and includes no discontinuities in its operations, thus being trainable with gradient-based methods.

\subsection{Architecture}
\paragraph{Overview}
An end-to-end memory network receives two inputs; a sequence of vectors $]\pmb{x}_1,\dots,\pmb{x}_N]$ and a query $\pmb{q}$. Both $\pmb{x}_i$ and $\pmb{q}$ are defined as symbols originating from a vocabulary $\mathcal{V}$. The network processes these inputs to return an answer $\pmb{a}$. This output is produced by first constructing continuous representations for both the input sequence and the query, and then iterating over these representations in many \textit{hops}, i.e. computational steps. The authors treat two different cases of this process; the first and simplest one employing a single layer, and the second, more general one, employing multiple layers. As this approach is favorable for understanding the architecture, we are also going to follow along the same route, modulo a few minor alterations and clarifications.

@@

As soon as we obtain the above, the matching vector $\pmb{p}$ between the query transformation and each memory object can be computed by passing the product of the two, $\pmb{u}^\top \pmb{m}_i$, through a Softmax activation:
\[
\tag{Memory Matching}
\pmb{p}_i = Softmax(\pmb{u}^\top \cdot \pmb{m}_i) = \frac{e^{\pmb{u}^\top\pmb{m}_i}}{\sum_j{e^{\pmb{u}^\top\pmb{m}_j}}}
\]
The inner product can be seen as a measure of similarity (reminiscent of \ref{eqn:Cosine}, minus the normalization), and the softmax activation is used to constrain $\pmb{p}$ within $\mathcal{S}^N$.

A third transformation, via a third embedding matrix $\pmb{C} \in \mathbb{R}^{d\times V}$, is used to map inputs $\{\pmb{x}_i\}$ to their output representations $\{\pmb{c}_i\}$, from which the memory response can be derived as their match-weighted sum:
\begin{align}
\tag{Output Transformation}
\pmb{c}_i &= \pmb{C}\pmb{x}_i\\
\tag{Memory Response}
\pmb{o} &= \sum\limits_i{\pmb{p}_i\pmb{c}_i}
\end{align}

@@

Finally, the memory response is combined with the query representation through some binary operation (e.g. concatenation, element-wise addition or multiplication, etc.) and is passed through one last layer responsible for the final prediction. In the original paper, which assumes a classification setting, the authors propose a Softmax-activated linear map over the addition of the query and the response:
\[
\tag{Final Prediction}
\hat{\pmb{a}} = Softmax(\pmb{W} (\pmb{o}+\pmb{u}))
\]

>>>>

\documentclass[]{article}
\usepackage[a4paper]{geometry}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathabx}
\usepackage{graphics}
%\usepackage{mathtools}
\usepackage{tikz}
\usepackage{array}
\usepackage{proof}
\usepackage{hyperref}
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour}
\usepackage{subcaption}
\usetikzlibrary{calc, trees, fit, positioning}
\usepackage{bm}

@@

A particularly interesting class of neural networks are \textit{Recurrent Neural Networks} (RNNs). Unlike fully-connected networks, which accept a fixed-size input vector and produce a fixed-size output vector over a finite number of computational steps, RNNs instead operate on sequences of vectors. This additional axis of dependencies is captured by a recurrency relation interlinking the network's hidden states through time. More concretely, the hidden representation is no longer conditioned on just the current input, but also on the immediately previous version of itself (Fig.\ref{fig:1}). This elegant addition accounts for a drastic improvement in the representational capacity of the network. Whereas a feedforward net acts as a trainable (but parametrically fixed) function, an RNN acts as a continuous program operating on sequences of inputs, and is in fact Turing complete\cite{Siegelmann545}. Under this viewpoint, the RNN's hidden representation can be perceived as the program's internal variables, updating as new inputs are received and affecting the computation as it progresses.

\begin{figure}
\begin{minipage}{0.5\textwidth}
	\centering
	\begin{tikzpicture}[scale=0.5]
		\node (X) at (-1, -2) {$\pmb{x}$};
		\foreach \y /\alph/\name in {
			0/x1/{},
		  	-1/x2/{},
		  	-2/x3/{},
		  	-4/xM/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=4mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (H) at (4, 2.5) {$\pmb{h}$};
		 \node (x..) at (0,-3) {\tiny {\vdots}};
		 \foreach \y /\alph/\name in {
			1.5/h1/{},
		  	0/h2/{},
		  	-1.5/h3/{},
		  	-5.5/hN/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (4,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \node (Y) at (9, -2) {$\pmb{y}$};
		 \foreach \y /\alph/\name in {
			0/y1/{},
		  	-1/y2/{},
		  	-2/y3/{},
		  	-4/yP/{}
		 }{
		 	\node[circle, draw=gray!90,inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
	\end{tikzpicture}
	\subcaption{Simple Neural Net}
	\end{minipage}
	\begin{minipage}{0.5\textwidth}
	\centering
	\begin{tikzpicture}[scale=.5]
		\node (X) at (-1, -2) {$\pmb{x}$};
		\foreach \y /\alph/\name in {
			0/x1/{},
		  	-1/x2/{},
		  	-2/x3/{},
		  	-4/xM/{}
		 }{
		% 	\node[] (\alph) at (0,\y) {\name};
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (0,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (x..) at (0,-3) {\tiny {\vdots}};
	  \node (H) at (4, 2.5) {$\pmb{h}$};
		 \foreach \y /\alph/\name in {
			1.5/h1/{},
		  	0/h2/{},
		  	-1.5/h3/{},
		  	-6.5/hN/{}
		 }{
		 	\node[circle, draw=black, inner sep=0pt,minimum size=3mm,fill=gray!20] (\alph) at (4,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (h..) at (4,-3) {\tiny {\vdots}};
		 \node (Y) at (9, -2) {$\pmb{y}$};
		 \foreach \y /\alph/\name in {
			0/y1/{},
		  	-1/y2/{},
		  	-2/y3/{},
		  	-4/yP/{}
		 }{
		 	\node[circle, draw=gray!90, inner sep=0pt,minimum size=3mm,fill=white] (\alph) at (8,\y) {\tiny \textcolor{black}{\name}};
		 }
		 \node (y..) at (8,-3) {\tiny {\vdots}};
		 \foreach \x in {x1, x2, x3, xM}
		 {\foreach \h in {h1, h2, h3, hN}
		 {\draw (\x) edge[gray!90] (\h);}}
		 \foreach \x in {h1, h2, h3, hN}
		 {\foreach \y in {y1, y2, y3, yP}
		 {\draw (\x) edge[gray!90] (\y);}}
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(0.8,0.8) and +(-0.8,0.8) .. (h1.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(0.4,-0.8) and +(-0.4,0.8) .. (h2.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(1,-1.6) and +(-0.6,0.4) .. (h3.west);
		 \draw (h1.east) [->, densely dotted, thick, black] .. controls +(2,-2) and +(-2,2) .. (hN.west);
\end{tikzpicture}
	\subcaption{Recurrent Neural Network}
	\end{minipage}
	\caption[]
	{\small \textit{Comparison between a feedforward neural network and an RNN. (a) The simple network applies two transformations $\pmb{x} \mapsto \pmb{h}$ and $\pmb{h} \mapsto \pmb{y}$, the composition of which can be seen as a function $\pmb{x} \mapsto \pmb{y}$. (b) The recurrent network allows through-time connections, depicted with the dotted lines, such that $\pmb{x}_t \times \pmb{h}_{t-1} \mapsto \pmb{h}_t$. The output then becomes a function of the full sequence of inputs received: $\underbrace{\pmb{x}_0 \times \pmb{x}_1 \times \dots \times \pmb{x}_t}_{\text{Sequence Length}} \mapsto \pmb{y}_t$.}\label{fig:1}}
\end{figure}

@@

A series of advancements has seeked to enhance RNNs with stronger memory management. Starting with the introduction of the \textit{Long Short-Term Memory} (LSTM)\cite{Hochreiter:1997:LSM:1246443.1246450,Gers99learningto}, which permits finite storage over very long computations, research in RNNs has bloomed towards many different directions. One particularly exciting recent advancement is the so-called \textit{attentional interface}\cite{NIPS2014_5542,Bahdanau} (Fig.\ref{fig:2}), which allows a network to selectively shift its focus on particular indices over a set of dimensions of an arbitrary tensor, thus enabling content-based adressing. The implications of this seemingly simple trick are tremendous; when employing fully differentiable attention (soft-attention), a network can learn how to pick out and combine relevant elements out of a (potentially very large) tensor, instead of lossily encoding it. This translates to arbitrary information retrieval; in other words, the same representations can now have different effects on the computation depending on the present context, similar to how a computer program would refer only to the variables dictated by its currently run instruction\cite{NIPS2014_5346}. Since its inception, attention has been a key-concept in many influential publications; the creative liberty it is offering, in terms of new potential operations, has also paved the way for architectures which decouple storage from computation. Such architectures, including (but not limited to) the two presented in this paper, utilize large external matrices and employ attention to perform read and write operations on them.

@@

		\draw (Ny1)  [dashed] .. controls +(0.8,0.8) and +(-0.8,0.8) ..  (x1);
		\draw (Ny2)  [dashed] .. controls +(3,3) and +(-0.6,0.6) ..  (x2);
		\draw (Ny3)  [dashed] .. controls +(2,-3) and +(0.6,0.4) ..  (xT.west);
	\end{tikzpicture}
	\caption[]
	{\small \textit{\label{fig:2} General case of an attentional layer. Given some sequence $[\pmb{h}_1,\dots,\pmb{h}_T]$ (e.g. a full sequence of RNN hidden states), we may use a neural function $a$ to obtain a weighting vector $\pmb{a} \in \mathbb{R}^{T}$. Multiplying each sequence element by its corresponding weight and summing up the results, we end up with a context vector $\pmb{c}$. Note that when $a$ accepts an external vector (or vector sequence) as its argument, the result is a context-dynamic selection and linear combination of (parts of) the $[\pmb{h}_1,\dots]$ sequence.}}
\end{figure}

@@

\section{Differentiable Neural Computers}
\label{DNC}
\subsection{Description}
Inspired by the von Neumann architecture, Differentiable Neural Computers (or DNCs in short) model computation as the interfacing between two units; a neural controller and an external memory matrix, functioning analogously to the CPU and the RAM, respectively. The memory matrix can be adressed using soft attention mechanisms, allowing selective reading and writing. As both the controller and the interfacing are differentiable, the full architecture is end-to-end differentiable, essentially allowing it to be trained purely with data and minimal human supervision. Since the system is already equiped with the necessary tools to simulate traditional computation, i.e. adressing mechanisms, information storage and retrieval, logical flows and arithmetic processing, a DNC in training is learning the instructions of a program rather than the parameters of a function. A trained DNC should then be able to mimic the behavior of a traditionally hand-written program.

@@

The retention vector is a measure of how much each memory address needs to be retained. Its functional complement, the \textit{usage-tracking vector} $\pmb{u}_t$, annotates each memory address with a scalar, responsible for signaling how strongly it is being used. It is recurrently defined as follows:
\[
\tag{Usage-tracking Vector}
\pmb{u}_t = \underbrace{(\pmb{u}_{t-1} + \pmb{w}_{t}^{w} - \pmb{u}_{t-1} \circ \pmb{w}_t^{w})}_{\text{ \scriptsize Previous usage OR New write}} \circ \ \pmb{\psi}_t
\]
Contrary to first impressions, the above equation is quite intuitive. The parenthesized term sums the previous usage tracking vector with the current write weighting minus their dot product, acting as a real-valued variation of the logical disjunction function. A single element of this term, $\pmb{u}_t[i]$ would denote the pre-retention usage tracking of memory address $i$; it is close to $1$ if it was either already in use, or just written to. Multiplied by its corresponding element of the retention vector, it can then be (partially) deallocated.

@@

\paragraph{Temporal Linking} One last mechanism of memory addressing employed by DNCs is temporal linking. Temporal linking is necessary in order to perform sequential retrieval, when sequences are stored in disparate, non-contiguous blocks. The key-element of the process is a \textit{temporal transition matrix} $\pmb{L} \in [0,1]^{N\times N}$, which keeps track of the temporal order by which memory addresses have been written to. More conretely, $\pmb{L}[i,j]$ is close to $1$ if $j$ was strongly written to after $i$, or close to $0$ otherwise. $\pmb{L}$ is again defined via a recurrency relation; a slightly simplified version of the one in the original paper is presented below, for the sake of brevity and clarity.
\[
\tag{Temporal Update}
\pmb{L}[i,j] = \underbrace{(1 - \pmb{w}_t^w[i] - \pmb{w}_t^w[j]) \pmb{L}_{t-1}[i,j]}_{\text{\scriptsize Part of last transition}} + \underbrace{\pmb{w}_{t-1}^w[i] \pmb{w}_{t}^w[j]}_{\text{\scriptsize New transition}}
\]
The above equation is also quite intuitive. The first term of the addition allows the transition value between $i$ and $j$ to propagate through time. The extend of propagation of the old transition, $\pmb{L}_{t-1}[i,j]$, is complentary to the amount of writing performed in these addresses; even if a transition from $i$ to $j$ had taken place in some past step, it no longer holds if either $i$ or $j$ were just written to. On the other hand, if $i$ was written to in the previous timestep and $j$ was written to in the current timestep, then a new transition is instantiated. This is captured by the second term of the addition.

@@

\section{End-to-End Memory Networks}
\label{E2E}
\subsection{Description}
An end-to-end memory network is another neural architecture that connects an RNN with a large memory array. The external nature of this array is disanalogous to traditional architectures, such as the LSTM, where the memory is encapsulated within the neural layers. Considering this difference, it can be perceived as true long-term memory. The memory can be read from (potentially multiple times) during computation, using a straightforward attentional layer; the architecture therefore benefits from context-sensitive addressing capabilities. The overall model is simple, yet efficient, and includes no discontinuities in its operations, thus being trainable with gradient-based methods.

\subsection{Architecture}
\paragraph{Overview}
An end-to-end memory network receives two inputs; a sequence of vectors $[\pmb{x}_1,\dots,\pmb{x}_N]$ and a query $\pmb{q}$. Both $\pmb{x}_i$ and $\pmb{q}$ are consist of symbols originating from a vocabulary $\mathcal{V}$. The network processes these inputs to return an answer $\pmb{a}$. This output is produced by first constructing continuous representations for both the input sequence and the query, and then iterating over these representations in many \textit{hops}, i.e. computational steps. The authors treat two different cases of this process; the first and simplest one employing a single layer, and the second, more general one, employing multiple layers. As this approach is favorable for understanding the architecture, this section follows along the same route, modulo a few minor alterations and clarifications.

@@

As soon as we obtain the above, the matching vector $\pmb{p}$ between the query transformation and each memory object can be computed by passing the product of the two, $\pmb{u}^\top \pmb{m}_i$, through a softmax activation:
\[
\tag{Memory Matching}
\pmb{p}_i = Softmax(\pmb{u}^\top \cdot \pmb{m}_i) = \frac{e^{\pmb{u}^\top\pmb{m}_i}}{\sum_j{e^{\pmb{u}^\top\pmb{m}_j}}}
\]
The inner product can be seen as a measure of similarity (reminiscent of \ref{eqn:Cosine}, minus the normalization), and the softmax activation is used to constrain $\pmb{p}$ within $\mathcal{S}^N$.

A third transformation, via an embedding matrix $\pmb{C} \in \mathbb{R}^{d\times V}$, is used to map inputs $\pmb{x}_i$ to their output representations $\pmb{c}_i$, from which the memory response can be derived as their match-weighted sum:
\begin{align}
\tag{Output Transformation}
\pmb{c}_i &= \pmb{C}\pmb{x}_i\\
\tag{Memory Response}
\pmb{o} &= \sum\limits_i{\pmb{p}_i\pmb{c}_i}
\end{align}

@@

Finally, the memory response is combined with the query representation through some binary operation (e.g. concatenation, element-wise addition or multiplication, etc.) and is passed through one last layer responsible for the final prediction. In the original paper, which assumes a classification setting, the authors propose a softmax-activated linear map over the addition of the query and the response:
\[
\tag{Final Prediction}
\hat{\pmb{a}} = Softmax(\pmb{W} (\pmb{o}+\pmb{u}))
\]
